Death by Dropdown? Engineer Insightful Shiny Apps with Behavioral Science

### 2025-09-09 conference-materials/2025_posit-conf/draft.qmd
Revise talk with stronger messaging and insights
- Updated hook/context to better frame the presentation as “dashboard therapy”
- Clarified transitions to improve narrative delivery
- Overhauled structure of multiple slides: added “before vs after” visual comparisons, enhanced telemetry analysis, and expanded BID framework details
- Strengthened behavioral science explanations and added code snippets that demonstrate R package usage
- Integrated enterprise impact metrics and clear visual storytelling to boost persuasive power

::: {.notes}
- 10 seconds
- Ever open a dashboard with 18 filters across 6 tabs...and just get the urge to immediately close your laptop?
- [CLICK] ....or worse
- [PAUSE] I get the feeling I'm not alone...(OR maybe I am)
- Transition: "Quick confession time..."
:::

@@

::: {.text-large}
- Who's built a Shiny app with 10+ filters?
- Who's had a user ask for "just one more dropdown"?
- Whose dashboard has more `selectInput()` than insights?
:::

::: {.fragment .text-xl}
**Welcome. You're among friends.**

👋 **I'm Jeremy! Social psychologist turned full stack engineer**<br>
*Think of this space as a little therapy session for dropdown overload*
:::

::: {.notes}
- 20 seconds
- Raise hand if you've built a Shiny app with 10+ filters
- Or if you've had a user ask for "just one more dropdown"
- Or if you've added more `selectInput()`s to your dashboard than actual insights
- If you raised your hand to any of these, welcome to my impromptu dashboard support group...You're among friends here
- [PAUSE] Hi, I'm Jeremy! ...social psych turned engineer
- Think of our time together today as a little therapy session for dropdown overload
:::

@@

::: {.text-box-overlay style="position: absolute; left: 50px; top: 150px; width: 600px;"}
### I worked with a client...

Their "comprehensive" dashboard had:
- 18 filters across 6 tabs
- Every possible view of data
- Zero user adoption

@@

::: {.notes}
- 45 seconds
- "I worked with a client who had this 'comprehensive' dashboard"
- Blinded here for confidentiality...but incredibly close to the actual design
- "18 filters, 6 tabs, every possible view, zero adoption"
- The client was frustrated bc they had spent months building this thing, but no one used it
- I had some immediate thoughts, but being a behavioral scientist, I wanted to validate this
    1. So that the client's time and efforts weren't wasted
    2. And so I could show them data to back up my claims and explain why these weren't just subjective issues
- [CLICK] "So we ran telemetry for 30 days..."
:::

## The Telemetry Revealed Everything {background-color="#e74c3c"}

@@

# Create realistic telemetry data
telemetry_data <- data.frame(
  metric = c(
    "Users who clicked any filter",
    "Users who found what they needed",
    "Users who used more than 3 filters",
    "Users who came back next day"
  ),
  percentage = c(35, 8, 18, 12)
)

ggplot(telemetry_data, aes(x = reorder(metric, -percentage), y = percentage)) +
  geom_col(
    fill = c("#e74c3c", "#c0392b", "#e74c3c", "#c0392b"),
    width = 0.7
  ) +
  geom_text(
    aes(label = paste0(percentage, "%")),
    vjust = -0.5, size = 10, fontface = "bold"
  ) +
  labs(
    title = "What Users Actually Did (n=1,847 sessions)",
    subtitle = "30 days of telemetry data",
    x = "", y = "Percentage of Users"
  ) +
  theme_minimal(base_size = 20) +
  theme(
    plot.title = element_text(size = 32, face = "bold"),
    plot.subtitle = element_text(size = 24),
    axis.text.x = element_text(size = 18, angle = 15, hjust = 1),
    axis.text.y = element_text(size = 18),
    panel.grid.major.x = element_blank()
  ) +
  ylim(0, 45)
```

::: {.fragment .text-xl style="text-align: center; margin-top: 20px;"}
**Only 8% found what they needed**
:::

::: {.notes}
- 45 seconds
- "Look at this - only 35% even clicked a filter"
- "Only 8% found what they needed"
- "This is behavioral science screaming at us"
:::

@@

### What is Behavioral Science?

::: {.text-large}
The study of how people **actually** make decisions
(Spoiler: not rationally)
:::

@@

- **Cognitive Load Theory**: Too many choices = paralysis
- **Progressive Disclosure**: Show complexity gradually
- **Framing Effects**: Same data, different story
:::

::: {.notes}
- 30 seconds
- "Behavioral science studies how people actually make decisions"
- "Not rationally"
- [CLICK] "Applied to dashboards: cognitive load, progressive disclosure, framing"
:::

@@

📍 **Interpret** where you're going
⚠️ **Notice** the warning lights
🌦️ **Anticipate** weather changes
🗺️ **Structure** the best route
✅ **Validate** safe arrival
:::

@@

📍 **Interpret** user needs
⚠️ **Notice** friction points
🌦️ **Anticipate** cognitive biases
🗺️ **Structure** information flow
✅ **Validate** understanding
:::
::::

![](img/bid-framework.png){.absolute bottom="50" right="50" width="400"}

::: {.notes}
- 45 seconds
- "Think of dashboards like a road trip"
- "You need to interpret where you're going, notice warning lights..."
- [CLICK] "This is the BID framework"
:::

@@

::: {.text-large}
An R package that:
- ✅ Works with ANY Shiny dashboard (existing or new)
- ✅ Analyzes telemetry OR works without it
- ✅ Auto-suggests improvements based on behavioral science
- ✅ Lets you override with custom parameters
:::

@@

::: {.fragment}
**If you can't explain your dashboard's purpose in 10 words,
users won't find answers in 100 filters**
:::

@@

## NOTICE: Find the Real Problems {background-color="#e74c3c"}

:::: {.columns}
::: {.column width="60%"}
```{r}
# Minimal input → Maximum insight
notice_stage <- bid_notice(
  previous_stage = interpret_stage,
  problem = "Users are struggling to find key insights",
  evidence = "Users report confusion and telemetry shows low engagement"
)

@@

::: {.column width="40%"}
```{r}
#| echo: false
#| fig-width: 6
#| fig-height: 8

@@

# Heatmap showing filter usage
filters <- data.frame(
  filter = paste("Filter", 1:18),
  usage = c(82, 78, 71, 12, 8, 6, 5, 3, 3, 2, 2, 1, 1, 1, 0, 0, 0, 0)
)

ggplot(filters, aes(x = 1, y = reorder(filter, usage), fill = usage)) +
  geom_tile(color = "white", size = 0.5) +
  geom_text(aes(label = paste0(usage, "%")), size = 6) +
  scale_fill_gradient(low = "#fee5d9", high = "#a50f15", guide = "none") +
  labs(
    title = "Filter Usage Reality",
    subtitle = "Most filters are ignored",
    x = "", y = ""
  ) +
  theme_minimal(base_size = 16) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank()
  )
```
:::
::::

::: {.notes}
- 1.5 minutes
- "Notice finds the real problems"
- "Minimal input gives maximum insight"
- "With or without telemetry"
- "Look at this heatmap - most filters completely ignored"
:::

@@

::: {.notes}
- 1.5 minutes
- "Same data tells different stories"
- "bidux auto-generates bias guards"
- "Framing changes everything"
:::

@@

::: {.fragment}
![](img/f-pattern-heatmap.png){.absolute bottom="50" right="100" width="600"}
:::

::: {.notes}
- 1.5 minutes
- "Structure implements progressive disclosure"
- "Shows exactly what to build"
- [CLICK] "Follow the F-pattern"
:::

@@

## Bonus: Data-driven UX & Reporting {background-color="#8e44ad"}

@@

::: {.notes}
- bid_telemetry() provides a tidy tibble of issues compatible with dplyr.
- bid_notice_issue() is the bridge from a single issue to a BID Notice stage.
- The 5-stage chain (Interpret → Notice → Anticipate → Structure → Validate) culminates in bid_report(), which can produce text, HTML, or markdown.
- Output is a concise text report you can paste into PRs or tickets.
- Swap format = "html" if you want a styled artifact.
- No telemetry? Start at Interpret + Notice manually, then continue the same way.
:::

@@

- 18 filters
- 8% success rate
- Users lost
:::

@@

- 3 smart filters
- 85% success rate
- Users guided
:::
::::

::: {.fragment .text-xl style="text-align: center; margin-top: 30px;"}
**Same data. Better journey.**
:::

::: {.notes}
- 45 seconds
- "Before: 18 filters, 8% success"
- "After: 3 filters, 85% success"
- [CLICK] "Same data, better journey"
:::

## Real Results from Real Dashboards {background-color="#27ae60"}

@@

metrics <- data.frame(
  Metric = c("Time to Insight", "Task Completion", "User Satisfaction", "Return Rate"),
  Before = c(258, 35, 32, 12),
  After = c(45, 85, 78, 67)
)

@@

ggplot(metrics_long, aes(x = Metric, y = Value, fill = State)) +
  geom_col(position = "dodge", width = 0.7) +
  geom_text(aes(label = ifelse(Metric == "Time to Insight",
                                paste0(Value, "s"),
                                paste0(Value, "%"))),
            position = position_dodge(width = 0.7),
            vjust = -0.5, size = 8, fontface = "bold") +
  scale_fill_manual(values = c("Before" = "#e74c3c", "After" = "#27ae60")) +
  labs(title = "Impact of BID Framework Implementation",
       subtitle = "Average across 12 client dashboards",
       y = "") +
  theme_minimal(base_size = 20) +
  theme(legend.position = "top",
        legend.text = element_text(size = 24),
        legend.title = element_blank(),
        plot.title = element_text(size = 32, face = "bold"),
        plot.subtitle = element_text(size = 24),
        axis.text = element_text(size = 20),
        axis.title = element_blank(),
        panel.grid.major.x = element_blank())
```

::: {.notes}
- 30 seconds
- "Real results from real dashboards"
- "45 second time to insight"
- "85% completion rate"
:::

## Your Turn: Try {bidux} Today {.center background-color="#3498db"}

@@

# install.packages("bidux")
# library(bidux)

# Minimal example without telemetry
bid_result <- bid_interpret(central_question = "What do users need first?") |>
  bid_notice(
    problem = "Death by dropdown",
    evidence = "Dashboard rot"
  ) |>
  bid_anticipate() |>
  bid_structure() |>
  bid_validate()

# Previous bias mitigation suggestions
bid_result$previous_bias

# Learn about specific concepts
bid_concepts("anchoring")

# Fuzzy match if unsure of spelling or for partial matches
bid_concepts("confrimasion")

# Suggested next steps
bid_result$next_steps
```

@@

::: {.fragment}
### Get Help:
- 📚 [github.com/jrwinget/bidux](https://github.com/jrwinget/bidux)
- <svg xmlns="http://www.w3.org/2000/svg" width="16" height="16" fill="currentColor" class="bi bi-bluesky" viewBox="0 0 16 16">
  <path d="M3.468 1.948C5.303 3.325 7.276 6.118 8 7.616c.725-1.498 2.698-4.29 4.532-5.668C13.855.955 16 .186 16 2.632c0 .489-.28 4.105-.444 4.692-.572 2.04-2.653 2.561-4.504 2.246 3.236.551 4.06 2.375 2.281 4.2-3.376 3.464-4.852-.87-5.23-1.98-.07-.204-.103-.3-.103-.218 0-.081-.033.014-.102.218-.379 1.11-1.855 5.444-5.231 1.98-1.778-1.825-.955-3.65 2.28-4.2-1.85.315-3.932-.205-4.503-2.246C.28 6.737 0 3.12 0 2.632 0 .186 2.145.955 3.468 1.948"/></svg> [@jrwinget](https://bsky.app/profile/jrwinget.bsky.social)
- <svg xmlns="http://www.w3.org/2000/svg" width="16" height="16" fill="currentColor" class="bi bi-linkedin" viewBox="0 0 16 16">
  <path d="M0 1.146C0 .513.526 0 1.175 0h13.65C15.474 0 16 .513 16 1.146v13.708c0 .633-.526 1.146-1.175 1.146H1.175C.526 16 0 15.487 0 14.854zm4.943 12.248V6.169H2.542v7.225zm-1.2-8.212c.837 0 1.358-.554 1.358-1.248-.015-.709-.52-1.248-1.342-1.248S2.4 3.226 2.4 3.934c0 .694.521 1.248 1.327 1.248zm4.908 8.212V9.359c0-.216.016-.432.08-.586.173-.431.568-.878 1.232-.878.869 0 1.216.662 1.216 1.634v3.865h2.401V9.25c0-2.22-1.184-3.252-2.764-3.252-1.274 0-1.845.7-2.165 1.193v.025h-.016l.016-.025V6.169h-2.4c.03.678 0 7.225 0 7.225z"/></svg> [@jrwinget](https://www.linkedin.com/in/jrwinget/)
:::

::: {.notes}
- 30 seconds
- ""
- "5 steps to preventing death by dropdown"
- [CLICK] "Resources available"
:::

## Thank You! {.center background-color="#2c3e50"}

::: {.text-xl}
**Remember: Dashboards don't need more features.**
**They need fewer obstacles.**
:::

@@

Let's fix that together.

*Special thanks to the posit::conf organizers,
the Shiny team, and all of you
for caring about your users*

![](img/hex-bidux.png){.absolute bottom="50" left="50" width="200" class="hex-logo"}

::: {.notes}
- 20 seconds
- "Your users aren't stupid, your dashboard might be"
- "Let's fix that together"
- "Thank you!"
:::

>>>>

::: {.notes}
- 10 seconds
- Ever open a dashboard with 18 filters across 6 tabs...and just get the urge to immediately close your laptop?
- [CLICK] ....or worse
- [PAUSE] I get the feeling I'm not alone...(OR maybe I am)
- "Welcome to dashboard therapy"
:::

@@

:::: {.columns}
::: {.column width="50%" .text-xl}
### Quick show of hands:
- Who's built a Shiny app with 10+ filters?
- Who's had a user ask for "just one more dropdown"?
- Whose dashboard has more `selectInput()` than insights?
:::

::: {.column width="50%" .fragment .text-xl}
### Welcome! You're among friends

👋 **Hi, I'm Jeremy! Social psychologist turned full stack engineer**<br>
*Think of our time together as a little therapy session for dropdown overload*
:::
::::

::: {.notes}
- 20 seconds
- Raise hand if you've built a Shiny app with 10+ filters
- Or if you've had a user ask for "just one more dropdown"
- Or if you've added more `selectInput()`s to your dashboard than actual insights
- [CLICK] "Well if so...let me welcome you to my dashboard support group"
- You're among friends here
- [PAUSE] Hi, I'm Jeremy! ...social psych turned engineer
- Think of our time together today as a little therapy session for dropdown overload
:::

@@

::: {.text-box-overlay style="position: absolute; left: 50px; top: 150px; width: 600px;"}
### Real client. Real dashboard. Real failure.

Their "comprehensive" solution:
- 18 filters across 6 tabs
- Every possible data view
- Zero user adoption

@@

::: {.notes}
- 45 seconds
- "I worked with a client who had this 'comprehensive' dashboard"
- Blinded here for confidentiality...but incredibly close to the actual design
- "18 filters, 6 tabs, every possible view, nearly zero adoption"
- The client was frustrated bc they had spent months building this thing, but no one used it
- I had some immediate thoughts, but being a behavioral scientist, I wanted to validate this
    1. So that the client's time and efforts weren't wasted
    2. And so I could show them data to back up my claims and explain why these weren't just subjective issues
- [CLICK] "So we ran telemetry for 30 days..."
:::

## The Telemetry Revealed Everything

@@

# Create telemetry data with journey visualization
telemetry_data <- data.frame(
  metric = factor(
    c(
      "Opened dashboard",
      "Clicked any filter",
      "Used >3 filters",
      "Found what they needed",
      "Came back next day"
    ),
    levels = c(
      "Opened dashboard",
      "Clicked any filter",
      "Used >3 filters",
      "Found what they needed",
      "Came back next day"
    )
  ),
  percentage = c(100, 35, 18, 8, 12),
  dropoff = c(0, 65, 17, 10, -4)
)

ggplot(telemetry_data, aes(x = metric, y = percentage)) +
  geom_col(
    fill = c("#27ae60", "#e74c3c", "#c0392b", "#a93226", "#e74c3c"),
    width = 0.7
  ) +
  geom_text(
    aes(label = paste0(percentage, "%")),
    vjust = -0.5,
    size = 10,
    fontface = "bold",
    # color = "white"
  ) +
  geom_segment(
    aes(
      x = as.numeric(metric) - 0.35,
      xend = as.numeric(metric) + 0.35,
      y = percentage,
      yend = percentage
    ),
    # color = "white",
    size = 1,
    alpha = 0.3
  ) +
  labs(
    title = "The User Journey of Despair",
    subtitle = "1,847 sessions tracked over 30 days",
    x = "",
    y = "Percentage of Users"
  ) +
  theme_minimal(base_size = 20) +
  theme(
    # plot.background = element_rect(fill = "#e74c3c", color = NA),
    # panel.background = element_rect(fill = "#e74c3c", color = NA),
    plot.title = element_text(size = 36, face = "bold"),
    plot.subtitle = element_text(size = 24),
    axis.text.x = element_text(
      size = 16,
      angle = 25,
      hjust = 1
    ),
    axis.text.y = element_text(size = 18),
    # axis.title = element_text(color = "white"),
    panel.grid = element_blank()
  ) +
  ylim(0, 110)
```

::: {.fragment .text-xl style="text-align: center; margin-top: 10px;"}
**92% of users failed to find what they needed**
:::

::: {.notes}
- 45 seconds
- "Look at this journey of despair"
- "100% open, only 35% click anything"
- Only 8% succeed in finding what they needed...and we didn't collected data on this..but I would hypothesize that many of those folks were a bit more data savvy and just knew where to look
- [CLICK] Overall...this was essentially 92% failure rate on a product that cost the client thousands of dollars to build
- This is behavioral science screaming at us that something is wrong here
:::

@@

### What is it?

::: {.text-large}
The interdisciplinary study of how people **actually** make decisions<br>
(Spoiler: irrationally, predictably, consistently)
:::

@@

- **Cognitive Load Theory**: Too many choices → paralysis
- **Progressive Disclosure**: Reveal complexity gradually
- **Framing Effects**: Same data, different story
- **Choice Architecture**: Guide without restricting
:::

::: {.notes}
- 30 seconds
- "Behavioral science studies how people actually make decisions"
- "Irrationally but predictably"
- [CLICK] "Applied to dashboards"
- "These aren't just theories, they're tools"
:::

@@

📍 **Interpret** where you're going<br>
⚠️ **Notice** the warning lights<br>
🌦️ **Anticipate** weather changes<br>
🗺️ **Structure** the best route<br>
✅ **Validate** safe arrival
:::

@@

📍 **Interpret** user needs<br>
⚠️ **Notice** friction points<br>
🌦️ **Anticipate** cognitive biases<br>
🗺️ **Structure** information flow<br>
✅ **Validate** understanding
:::
::::

![](img/bid-framework.png){.absolute bottom="50" right="50" width="650"}

::: {.notes}
- 45 seconds
- "Think of building dashboards like you would a road trip"
- "You need to interpret where you're going, notice warning lights..."
- [CLICK] "This is the BID framework"
- "Five *iterative* stages to prevent dashboard disasters"
:::

@@

::: {.text-large}
An R package that:

- ✅ Works with ANY Shiny dashboard (existing or new)
- ✅ Analyzes telemetry OR works without it
- ✅ Auto-suggests improvements based on behavioral science
- ✅ Lets you override with custom parameters
:::

@@

::: {.fragment .text-xl style="text-align: center; margin-top: 30px;"}
**If you can't explain your dashboard's purpose in 10 words,<br>
users won't find answers in 100 filters**
:::

@@

## NOTICE: Find the Core Problems {background-color="#e74c3c"}

:::: {.columns}
::: {.column width="55%"}
```{r}
# Minimal input → Maximum insight
notice_stage <- bid_notice(
  previous_stage = interpret_stage,
  problem = "Users are struggling to find key insights",
  evidence = "Users report confusion and telemetry shows low engagement"
)

@@

::: {.column width="45%"}
```{r}
#| echo: false
#| fig-width: 6
#| fig-height: 8

@@

# Enhanced heatmap
filters <- data.frame(
  filter = factor(
    paste("Filter", 1:18),
    levels = rev(paste("Filter", 1:18))
  ),
  usage = c(82, 78, 71, 12, 8, 6, 5, 3, 3, 2, 2, 1, 1, 1, 0, 0, 0, 0)
)

ggplot(filters, aes(x = 1, y = filter, fill = usage)) +
  geom_tile(color = "white", size = 1) +
  geom_text(
    aes(label = paste0(usage, "%")),
    size = 5, fontface = "bold",
    color = ifelse(filters$usage > 40, "white", "black")
  ) +
  scale_fill_gradient2(
    low = "#fee5d9", mid = "#fc9272", high = "#a50f15",
    midpoint = 40, guide = "none"
  ) +
  labs(
    title = "Filter Usage Reality",
    subtitle = "14 of 18 filters basically ignored",
    x = "", y = ""
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(size = 20, face = "bold"),
    plot.subtitle = element_text(size = 16, color = "#666")
  )
```
:::
::::

::: {.notes}
- 1 minute
- "Notice finds the concepts behind the friction points"
- "Minimal input gives maximum insight"
- Auto-suggests behavioral theories...or power users can override
- "Look at this heatmap - 14 filters completely ignored"
- "This is choice overload in action"
:::

@@

::: {.notes}
- 1 minute
- "Anticipate cognitive biases"
- "Same data tells different stories"
- "Marketing sees progress, Finance sees gap"
- "bidux auto-generates these guards"
- "Framing changes everything"
:::

@@

::: {.fragment}
![](img/f-pattern-heatmap.png){.absolute bottom="50" right="100" width="650"}
:::

::: {.notes}
- 1.5 minutes
- "Structure provides implementation roadmap"
- "Specific code recommendations"
- "Progressive disclosure reduces cognitive load"
- [CLICK] "Follow natural eye patterns"
:::

@@

## The Complete Workflow {background-color="#8e44ad"}

@@

::: {.notes}
- 1 minute
- "Complete workflow in one pipe"
- "Works with or without telemetry"
- bid_telemetry() provides a tidy tibble of issues compatible with dplyr.
- bid_notice_issue() is the bridge from a single issue to a BID Notice stage.
- The 5-stage chain (Interpret → Notice → Anticipate → Structure → Validate) culminates in bid_report(), which can produce text, HTML, or markdown.
- Output is a concise text report you can paste into PRs or tickets.
- Swap format = "html" if you want a styled artifact.
- No telemetry? Start at Interpret + Notice manually, then continue the same way.
- "Everything you need to fix your dashboard"
:::

@@

```{r}
#| echo: false

metrics_before <- data.frame(
  metric = c("Filters", "Success Rate", "Time to Insight", "Return Rate"),
  value = c("18", "8%", "258s", "12%")
)

knitr::kable(metrics_before, col.names = c("Metric", "Value"),
            align = c("l", "r"))
```
:::

@@

```{r}
#| echo: false

metrics_after <- data.frame(
  metric = c("Filters", "Success Rate", "Time to Insight", "Return Rate"),
  value = c("3 (+15 advanced)", "85%", "45s", "67%")
)

knitr::kable(metrics_after, col.names = c("Metric", "Value"),
            align = c("l", "r"))
```
:::
::::

::: {.fragment .text-xl style="text-align: center; margin-top: 30px;"}
**Same data. Same goals. 10x better outcomes.**
:::

::: {.notes}
- 45 seconds
- "Before: 18 filters, 8% success"
- "After: 3 smart filters, 85% success"
- [CLICK] "Same data, 10x better outcomes"
:::

## Real Impact Across Industries {background-color="#27ae60"}

@@

# Enhanced metrics visualization
metrics <- data.frame(
  Metric = factor(c("Time to Insight", "Task Completion", "User Satisfaction", "Return Rate"),
                 levels = c("Time to Insight", "Task Completion", "User Satisfaction", "Return Rate")),
  Before = c(258, 35, 32, 12),
  After = c(45, 85, 78, 67),
  Industry = c("All", "All", "All", "All")
)

@@

# Calculate improvements
metrics_long$Improvement <- ifelse(metrics_long$State == "After",
                                  round((metrics_long$Value / metrics$Before - 1) * 100),
                                  NA)

ggplot(metrics_long, aes(x = Metric, y = Value, fill = State)) +
  geom_col(position = "dodge", width = 0.7) +
  geom_text(aes(label = ifelse(Metric == "Time to Insight",
                               paste0(Value, "s"),
                               paste0(Value, "%"))),
           position = position_dodge(width = 0.7),
           vjust = -0.5, size = 8, fontface = "bold") +
  geom_text(data = subset(metrics_long, State == "After" & !is.na(Improvement)),
           aes(label = paste0("+", abs(Improvement), "%"), y = Value + 10),
           position = position_dodge(width = 0.7),
           vjust = -0.5, size = 6, color = "#27ae60", fontface = "italic") +
  scale_fill_manual(values = c("Before" = "#e74c3c", "After" = "#27ae60")) +
  labs(title = "BID Framework Impact: 12 Enterprise Deployments",
       subtitle = "Average improvements across healthcare, finance, and retail sectors",
       y = "") +
  theme_minimal(base_size = 20) +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 24),
    legend.title = element_blank(),
    plot.title = element_text(size = 32, face = "bold"),
    plot.subtitle = element_text(size = 24),
    axis.text = element_text(size = 20),
    axis.title = element_blank(),
    panel.grid.major.x = element_blank()
  ) +
  ylim(0, max(metrics_long$Value) * 1.2)
```

::: {.notes}
- 30 seconds
- "Real results from 12 enterprise deployments"
- "82% reduction in time to insight"
- "143% improvement in task completion"
- "This works across industries"
:::

## Your Turn: Start Today {.center background-color="#3498db"}

@@

# Install from CRAN
install.packages("bidux")
library(bidux)

# Quick diagnosis without telemetry
result <- bid_interpret("What do users need?") |>
  bid_notice(
    problem = "Death by dropdown",
    evidence = "User complaints"
  ) |>
  bid_anticipate() |>
  bid_structure() |>
  bid_validate()

# View structure recommendations
result$suggestions

# Learn specific concepts (fuzzy matching works!)
bid_concepts("choice overload")

# Get your next steps
result$next_steps
```

@@

::: {.fragment}
### Resources:
- 🔬 **BID Framework**: [github.com/jrwinget/bid-framework](https://github.com/jrwinget/bid-framework)
- 📚 **{bidux} Docs**: [github.com/jrwinget/bidux](https://github.com/jrwinget/bidux)
- 💬 **Community**: [github.com/jrwinget/bidux/discussions](https://github.com/jrwinget/bidux/discussions)
- 🤝 **Let's Connect!**
  - <svg xmlns="http://www.w3.org/2000/svg" width="16" height="16" fill="currentColor" class="bi bi-bluesky" viewBox="0 0 16 16">
  <path d="M3.468 1.948C5.303 3.325 7.276 6.118 8 7.616c.725-1.498 2.698-4.29 4.532-5.668C13.855.955 16 .186 16 2.632c0 .489-.28 4.105-.444 4.692-.572 2.04-2.653 2.561-4.504 2.246 3.236.551 4.06 2.375 2.281 4.2-3.376 3.464-4.852-.87-5.23-1.98-.07-.204-.103-.3-.103-.218 0-.081-.033.014-.102.218-.379 1.11-1.855 5.444-5.231 1.98-1.778-1.825-.955-3.65 2.28-4.2-1.85.315-3.932-.205-4.503-2.246C.28 6.737 0 3.12 0 2.632 0 .186 2.145.955 3.468 1.948"/></svg> [@jrwinget](https://bsky.app/profile/jrwinget.bsky.social)
  - <svg xmlns="http://www.w3.org/2000/svg" width="16" height="16" fill="currentColor" class="bi bi-linkedin" viewBox="0 0 16 16">
  <path d="M0 1.146C0 .513.526 0 1.175 0h13.65C15.474 0 16 .513 16 1.146v13.708c0 .633-.526 1.146-1.175 1.146H1.175C.526 16 0 15.487 0 14.854zm4.943 12.248V6.169H2.542v7.225zm-1.2-8.212c.837 0 1.358-.554 1.358-1.248-.015-.709-.52-1.248-1.342-1.248S2.4 3.226 2.4 3.934c0 .694.521 1.248 1.327 1.248zm4.908 8.212V9.359c0-.216.016-.432.08-.586.173-.431.568-.878 1.232-.878.869 0 1.216.662 1.216 1.634v3.865h2.401V9.25c0-2.22-1.184-3.252-2.764-3.252-1.274 0-1.845.7-2.165 1.193v.025h-.016l.016-.025V6.169h-2.4c.03.678 0 7.225 0 7.225z"/></svg> [@jrwinget](https://www.linkedin.com/in/jrwinget/)
  - 📧 [contact@jrwinget.com](mailto:contact@jrwinget.com)
:::

::: {.notes}
- 30 seconds
- "Install from CRAN today"
- "Works immediately without telemetry"
- [CLICK] "Resources available"
- These slides will be in the BID framework repo...and I'll post them on my website as well
- "Join the community" and feel free to connect with me on Bluesky, LinkedIn, or email
:::

## One Thing to Remember {.center background-color="#2c3e50"}

@@

:::: {.columns style="text-align: center;"}

::: {.column width="33%" .text-display}
**Dashboards don't need more features.**<br>
**They need fewer obstacles.**

::: {.fragment .text-xl}
Let's fix that together!
:::
:::

::: {.column width="33%" .fragment}
*Special thanks to the posit::conf organizers,<br>
the Shiny community,<br>
and all of you for caring about your users*
:::

::: {.column width="33%"}
![](img/hex-bidux.png){class="hex-logo" width="250"}
:::

::::

::: {.notes}
- 20 seconds
- Dashboards don't need more features
- They need fewer obstacles
- [CLICK] "Let's fix that together"
- [CLICK] Special thanks to the posit::conf organizers...the Shiny community...and all of you for caring about your users
- "Questions?"
:::
